A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-20 paper/main.tex
finished first round of edits on methods, snapshot -> redshift focus, and updated plots/captions/description of the sample

%%% OUTLINE
\section{Methodology}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-XX$.

@@

\subsection{Pair orbit sample}
We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.

@@

The base sample used for this analysis then consists of the set of all low-mass and high-mass major pairs from each snapshot of the TNG100 simulation.

\subsubsection{Mergers}
% Plot \#1: Merger fraction either before or after orbits
% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).

The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each snapshot. 
The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.

We utilize the \texttt{RootDescendantID} and \texttt{DescendantID} fields of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$), and to determine when the merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshot, if it has one. 
% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.  
The \texttt{RootDescendantID} field provides the SubhaloID of the root subhalo, which is the latest subhalo in the simulation that is a descendent of the subhalo.

We define the set of merging pairs as the collection of all pairs where the primary and secondary halo have the same \texttt{RootDescendantID}, meaning that the branches of their merger trees have the same `root', or common descendent. 
If a given pair has two distinct \texttt{RootDescendantID} values, the pair is defined as a `non-merger.'

For each merging pair, we define the `merger snapshot' as the snapshot that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
For example, if the primary and secondary have different \texttt{DescendantIDs} for snapshots 40-45, but have the same \texttt{DescendantID} at snapshot 46, then the merger must take place some time between snapshot 46 and 47. 
We take snapshot 47 to be the merger snapshot.

\subsubsection{Orbits} 
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. 
    (Top) Orbits of pairs that merge before $z=0$.
    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
    Solid lines represent the parts of the orbit post-infall, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
    Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. \kc{state in words that the FoF group need not be common for the remainder of the orbit, but will definitely be by merger.}
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

We extract orbits for all mergers and non-mergers in our base sample. 
An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or lookback time.

A given pair from the base sample at snapshot $S_n$, which by definition passes all selection criteria at $S_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.  
We track the positions of both the primary and secondary subhalo at each snapshot and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
In cases where the primary or secondary does not have a defined position at a given snapshot, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs. The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be computed contiguously for a pair on . See Sec.~3 in ~\citet{RG2015} for more details.}

In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, given as
\begin{equation}
    \rm \Rsc = \frac{\rsep}{\Rvir}
\end{equation}
where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
The scaled radius $\Rsc$ is by construction a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs, or for halo growth over time.

We collect the following data for each pair at each snapshot where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).

\subsubsection{"First Infall"}
The orbit of a pair may be technically defined at very early times, well before the primary and secondary are physically associated. 
However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group. Specifically, we will consider only the ``Post-Infall" part of each pair's orbit.

We define the snapshot of``First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter. 
The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same. 
This is because there are cases of pair interactions where the secondary will enter the FoF group of the primary, have a pericentric passage, then leave the FoF group for a few gigayears before re-entering the primary FoF group and eventually merging with the primary.

An example of the orbit of a merging pair where a "common group" criteria may differ from a "post-infall" can be seen in Figure~\ref{fig:example-orbits}, which shows the variety of orbits that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of 5 pairs that merge, where solid lines show the parts of the orbit that occur while the primary and secondary share a common FoF group, and dashed lines show the parts of the merger where the FoF group of the secondary is different than that of the primary. 
The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
The spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
% a variety of different orbits

The middle panel shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a couple Gyr past the end of the simulation, and are thus "Potential Future Mergers". 
These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.

\subsubsection{Uniqueness of orbits}
Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.

To distinguish pairs from one snapshot to another, each pair is assigned a pairkey while constructing their orbit, which is created by combining the primary's and secondary's earliest SubhaloID from the Sublink catalogs, which will be unique for each pair of halos. 
After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.

For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
\begin{equation*}
    \sum_{k=0}^{99}\, n_k
\end{equation*}
The total number of low-mass pairs is 71,429, and the total number of high-mass pairs is 20,824.
After selecting only unique pairs, there remain 22,213 unique low-mass pairs, and 3,029 unique high-mass pairs.

Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are kept, since having the primary occur twice in the catalog does not mean that the orbit is being double counted. ugh.

Maybe need to define the "unique pair sample" and the "snapshot selected sample" to explain the plots? I think all plots EXCEPT the merger fraction are with unique pair sample.... DO I need to change that plot? 
% \subsection{Calculating the Number of Unique Pairs}

% \subsection{Calculating the Merger Fraction}

% \subsection{Calculating the Time Until Merger}

\section{Pair Sample Properties}
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{\kc{need caption}\kc{Change redshift limits?} The number of low-mass (green) and high-mass (pink) pairs as a function of redshift. The solid lines show the number of unique pairs at a given redshift, while dashed lines show the number of unique pairs at a given redshift that will merge before $z=0$.
    For these two lines, we have only counted pairs at a given redshift if they have not merged, but have experienced first infall at a previous time. 
    \kc{I believe... think more} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
    %
    The thin dotted lines show the number of pairs that were selected via the pair criteria at that specific redshift. 
    }
    \label{fig:num-pairs}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
    \caption{\kc{redshift:  0-6} The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}. At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.\kc{might need to add on second set of lines, so i have both merger fraction from snapshot-selected and from orbit-selected.}}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Unique Pairs}
We calculate the number of unique pairs at a given snapshot by selecting all unique pairs at that snapshot that have already experienced first infall, and then sum the number of pairs that have yet to merge.
Thus, the total number of unique pairs at $z=2$ includes pairs that were added to the catalog at \textit{any} snapshot, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (for example, a pair that was first selected in the snapshot-selected catalogs at $z=1$) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of unique pairs at all redshifts from $z=0-3$.

Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and low redshift selected pairs will on average have merger timescales longer than the remaining time to the end of the simulation.

The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met. 
These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.

\subsection{Merger Fraction}
We calculate the merger fraction, $\rm F_{merge}$, at a given snapshot by dividing the number of unique merging pairs (dashed lines in Fig.~\ref{fig:num-pairs}) by the number of total unique pairs (solid lines). 
As before, a pair whose first infall occurs at $z=2$ and that merges at $z=1$ will be included in the merger fraction calculation for redshift $z=1-2$.

Note that the merger fraction is a measure of the fraction of isolated pairs, whose primary and secondary subhalos have shared a common FoF group, that merge before $z=0$. 
It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$, but declines to $0$ at $z=0$. 
This is due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.

@@

** 
\appendix
adfadfa
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
%     \caption{}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
% \end{figure}

>>>>

%%% OUTLINE
\section{Methodology}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

@@

\subsection{Pair sample}
We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.

@@

The base sample of pairs used for this analysis then consists of the set of all low-mass and high-mass major pairs from each redshift of the TNG100 simulation.

\subsection{Mergers}
% Plot \#1: Merger fraction either before or after orbits
% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).

The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
The \texttt{SUBLINK} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.

We utilize the \texttt{DescendantID} field of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshot, if it has one. 
% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.

A pair will be classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
This means that the primary and secondary subhalo will merge, producing a single descendant subhalo, by the following redshift from the simulation.
If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'

For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift in the simulation. 
We take $z=1.9$ to be the merger redshift.

\subsection{Orbits} 
\begin{figure}[tb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. 
    (Top) Orbits of pairs that merge before $z=0$.
    (Middle) Orbits of pairs that do not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
    (Bottom) Orbits of pairs that do not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
    Solid lines represent the post-infall parts of the orbit, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
    Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. 
    }
    \label{fig:example-orbits}
    \end{center}
\end{figure}

We extract orbits for all mergers and non-mergers in our base pair sample. 
An orbit for a single pair is defined to be the physical separation between the primary and secondary subhalo as a function of redshift (or lookback time).

A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.  
We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
In cases where the primary or secondary does not have a defined position at a given redshift, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs. The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs. See Sec.~3 in ~\citet{RG2015} for more details.}

In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction of the virial radius $\rsep=X\Rvir$
% \begin{equation}
%     \rm \Rsc = \frac{\rsep}{\Rvir}
% \end{equation}
where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled radius" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

We collect the following data for each pair at each redshift where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).

\subsubsection{Defining Post-Infall}
While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group. Specifically, we will consider only the ``post-infall" part of each pair's orbit.

We define the redshift of ``first infall" as the redshift of the first snapshot where the primary and secondary have the same parent FoF halo, and post-infall is all following redshifts starting with the redshift of first infall.

Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of 5 pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo. 
Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\Gyr$ from first infall to merger.

Some secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo. 
Because all of our mergers occur when the subhalos are in the same FoF group, the
post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same. 
% a variety of different orbits

The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation. 
These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past $2\,\Gyr$.

The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.

\subsubsection{Uniqueness of orbits}
A given pair from the \paircat{} may pass our selection criteria at many redshifts. 
Since we collect the full orbit (the separation at every redshift where the primary and secondary both exist) for each pair and each redshift, the orbits that are collected for pairs that pass our selection criteria at more than one redshift will all result in identical orbits. 
% However, we only want to keep a single instance of any given orbit in our catalog.

To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.

Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.

Removing multiplicity decreases the raw number of orbits collected by collecting an orbit for each pair from the base sample. 
The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits.

% For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
% \begin{equation*}
%     \sum_{k=0}^{99}\, n_k
% \end{equation*}
% The total number, and the total number of high-mass pairs is .
% After selecting only unique pairs, there remain

The collection of all unique orbits constitutes our orbit sample, and is the dataset that will be used for the remainder of the paper.

\section{Pair Sample Properties}
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
    The solid lines show the total number of orbits the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
    We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
    \caption{The fraction of low-mass (green) and high-mass (pink) post-infall orbits that result in a merger before z=0 as a function of redshift. 
    The horizontal line shows a merger fraction of 0.9, or 90\%. 
    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
    The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
    \label{fig:fmerge}
    \end{center}
\end{figure}

\subsection{Number of Orbits}
We calculate the number of orbits (or orbits in progress) by selecting all orbits at a given redshift that are post-infall and either pre-merger or non-mergers.
The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (or equivalently, the pair's orbit) orbit may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of orbits at all redshifts from $z=0-3$.

Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since 
%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.

% removed
% The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met. 
% These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.

\subsection{Merger Fraction}
We calculate the merger fraction by dividing the number of merging orbits (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$. At $z=0$, any remaining pairs are classified as non-mergers since they did not merge \textit{$z=0$}.

Note that the merger fraction is a measure of the fraction of isolated pairs that merge before $z=0$. 
It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.

@@

% ** 
% \appendix
% adfadfa
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
%     \caption{}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
% \end{figure}
